---
title: "officer options"
author: "Andrew Ba Tran"
date: "May 26, 2016"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache=TRUE)
```

#Exploring different ways to visualize traffic data for individual police officers.

```{r cars, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggalt)
explore <- read.csv("data/mega_df.csv")
```

---

```{r pressure, fig.width=8, fig.height=5, echo=FALSE}
single_dept <- subset(explore, DepartmentName=="Waterbury" & is.na(ReportingOfficerIdentificationID))
single_officer <- subset(explore, DepartmentName=="Waterbury" & ReportingOfficerIdentificationID=="Waterbury--2103")
single_all <- subset(explore, DepartmentName=="Connecticut average"  & is.na(ReportingOfficerIdentificationID))

dept_name <- paste0(single_dept$DepartmentName[1], " Police | Traffic Stops")
```

## `r paste0("Officer: ", single_officer$ReportingOfficerIdentificationID[1])`


## `r paste0(single_dept$DepartmentName[1], " Police | Traffic Stops")`

### `r paste0(single_officer$min_p[1], "  percent of drivers stopped by Officer ", single_officer$ReportingOfficerIdentificationID[1], " were minorities, compared with ", single_dept$DepartmentName[1], "'s ", single_dept$min_p[1], " percent and state's ", single_all$min_p[1], " percent statewide.")`


```{r section1, fig.width=8, fig.height=4, echo=FALSE}

stops_race1 <- single_dept %>%
  select(DepartmentName, asian_p, black_p, hispanic_p, indian_p, mid_e_p, white_p) %>%
  gather("Ethnicity", "Percent", 2:7)

stops_race2 <- single_all %>%
  select(DepartmentName, asian_p, black_p, hispanic_p, indian_p, mid_e_p, white_p) %>%
  gather("Ethnicity", "Percent", 2:7)  

stops_race2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_race2$DepartmentName)


stops_race3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, asian_p, black_p, hispanic_p, indian_p, mid_e_p, white_p) %>%
  gather("Ethnicity", "Percent", 2:7)


names(stops_race3)[names(stops_race3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'


stops_race <- rbind(stops_race1, stops_race2)
stops_race <- rbind(stops_race, stops_race3)





stops_race$Ethnicity <- gsub("asian_p", "Asian", stops_race$Ethnicity)
stops_race$Ethnicity <- gsub("black_p", "Black", stops_race$Ethnicity)
stops_race$Ethnicity <- gsub("hispanic_p", "Hispanic", stops_race$Ethnicity)
stops_race$Ethnicity <- gsub("indian_p", "Indian", stops_race$Ethnicity)
stops_race$Ethnicity <- gsub("mid_e_p", "Middle-Eastern", stops_race$Ethnicity)
stops_race$Ethnicity <- gsub("white_p", "White", stops_race$Ethnicity)

ggplot(stops_race, aes(Ethnicity, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Traffic stops in 2014 by race percent")

#ggplot(stops_race, aes(x=DepartmentName, y=Percent)) + geom_bar(aes(fill=Ethnicity), stat="identity") +
#  geom_text(aes(label=Percent, y=pos), size=3) + coord_flip() + labs(title="Traffic stops in 2014")

```


### `r paste0(single_dept$minorities_p[1], " % NON-WHITE")`

The driving-age population is `r single_dept$minorities_p[1]` percent minority in `r single_dept$DepartmentName[1]` and `r single_all$minorities_p[1]` minority statewide.

---

## Gender

For officer `r single_officer$ReportingOfficerIdentificationID[1]`

* **`r single_officer$male_p[1]`** percent were men
* **`r single_officer$female_p[1]`** were women


In `r single_dept$DepartmentName[1]`

* **`r single_dept$male_p[1]`** percent were men
* **`r single_dept$female_p[1]`** were women

Statewide

* **`r single_all$male_p[1]`** percent were men
* **`r single_all$female_p[1]`** were women

---


### The median age for drivers stopped by officer `r single_officer$ReportingOfficerIdentificationID[1]` was `r single_officer$median_age[1]` versus `r single_dept$median_age[1]` in `r single_dept$DepartmentName[1]` and `r single_all$median_age[1]` years old statewide.


```{r section2, fig.width=8, fig.height=4, echo=FALSE}

stops_age1 <- single_dept %>%
  select(DepartmentName, y16_21_p, y22_27_p, y28_31_p, y32_37_p, y38_41_p, y42_47_p, y48_51_p, y52_57_p, y58_61_p, y62_67_p, y68_71_p, y72_77_p, y78_81_p, y82_p) %>%
  gather("Age", "Percent", 2:15)  

stops_age2 <- single_all %>%
  select(DepartmentName, y16_21_p, y22_27_p, y28_31_p, y32_37_p, y38_41_p, y42_47_p, y48_51_p, y52_57_p, y58_61_p, y62_67_p, y68_71_p, y72_77_p, y78_81_p, y82_p) %>%
  gather("Age", "Percent", 2:15) 

stops_age2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_age2$DepartmentName)

stops_age3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, y16_21_p, y22_27_p, y28_31_p, y32_37_p, y38_41_p, y42_47_p, y48_51_p, y52_57_p, y58_61_p, y62_67_p, y68_71_p, y72_77_p, y78_81_p, y82_p) %>%
  gather("Age", "Percent", 2:15) 

names(stops_age3)[names(stops_age3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'


stops_age <- rbind(stops_age1, stops_age2)
stops_age <- rbind(stops_age, stops_age3)


stops_age$Age <- gsub("y", "", stops_age$Age)
stops_age$Age <- gsub("_", "-", stops_age$Age)
stops_age$Age <- gsub("-p", "", stops_age$Age)
stops_age$Age <- gsub("82", "82+", stops_age$Age)

ggplot(stops_age, aes(Age, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Traffic stops in 2014 by age group percent")

```



---

### When

```{r section5, fig.width=8, fig.height=4, echo=FALSE}

stops_day1 <- single_officer %>%
  select(ReportingOfficerIdentificationID, Sun, Mon, Tues, Wed, Thurs, Fri, Sat) %>%
  gather("Day", "Percent", 2:8)  

stops_day <- stops_day1
stops_day$Day <- factor(stops_day$Day, levels= c("Sun", "Mon", 
    "Tues", "Wed", "Thurs", "Fri", "Sat"))

ggplot(stops_day, aes(Day, Percent, fill = ReportingOfficerIdentificationID, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Days when drivers were stopped") 

```


```{r section3, fig.width=6, fig.height=7, echo=FALSE}

stops_reason1 <- single_dept %>%
  select(DepartmentName, DefectiveLights_p, DisplayofPlates_p, EquipmentViolation_p, MovingViolation_p, Other_p, Registration_p, Seatbelt_p, SpeedRelated_p, StopSign_p, SuspendedLicense_p,TrafficControlSignal_p,
WindowTint_p) %>%
  gather("Reason", "Percent", 2:13)  

stops_reason2 <- single_all %>%  
  select(DepartmentName, DefectiveLights_p, DisplayofPlates_p, EquipmentViolation_p, MovingViolation_p, Other_p, Registration_p, Seatbelt_p, SpeedRelated_p, StopSign_p, SuspendedLicense_p,TrafficControlSignal_p,
WindowTint_p) %>%
  gather("Reason", "Percent", 2:13)  

stops_reason2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_reason2$DepartmentName)


stops_reason3 <- single_officer %>%  
  select(ReportingOfficerIdentificationID, DefectiveLights_p, DisplayofPlates_p, EquipmentViolation_p, MovingViolation_p, Other_p, Registration_p, Seatbelt_p, SpeedRelated_p, StopSign_p, SuspendedLicense_p,TrafficControlSignal_p,
WindowTint_p) %>%
  gather("Reason", "Percent", 2:13)  


names(stops_reason3)[names(stops_reason3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'


stops_reason <- rbind(stops_reason1, stops_reason2)
stops_reason <- rbind(stops_reason, stops_reason3)

stops_reason$Reason <- gsub("_p", "", stops_reason$Reason)

ggplot(stops_reason, aes(Reason, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), hjust = -0.2, size=3) + labs(title="Reasons for traffic stops") + coord_flip()

```



---

### After traffic stops

```{r section6, fig.width=8, fig.height=4, echo=FALSE}

stops_result1 <- single_dept %>%
  select(DepartmentName, infraction_ticket_p, misdemeanor_summons_p, no_disposition_p, uniform_arrest_p, verbal_warning_p, written_warning_p) %>%
  gather("Result", "Percent", 2:7)  


stops_result2 <- single_all %>%
  select(DepartmentName, infraction_ticket_p, misdemeanor_summons_p, no_disposition_p, uniform_arrest_p, verbal_warning_p, written_warning_p) %>%
  gather("Result", "Percent", 2:7)  

stops_result2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_result2$DepartmentName)


stops_result3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, infraction_ticket_p, misdemeanor_summons_p, no_disposition_p, uniform_arrest_p, verbal_warning_p, written_warning_p) %>%
  gather("Result", "Percent", 2:7)  

names(stops_result3)[names(stops_result3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'


stops_result <- rbind(stops_result1, stops_result2)
stops_result <- rbind(stops_result, stops_result3)

stops_result$Result <- gsub("_p", "", stops_result$Result)


ggplot(stops_result, aes(Result, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Result after traffic stops") 

```



---

### Searches

`r single_dept$DepartmentName[1]` officers searched `r single_dept$searched_p[1]` percent of drivers they stopped. `r single_officer$ReportingOfficerIdentificationID[1]` searched `r single_officer$searched_p[1]` percent.

This chart shows the percentage of all stops involving searches of cars driven by white or non-white drivers.

```{r section4, fig.width=8, fig.height=4, echo=FALSE}

stops_search1 <- single_dept %>%
  select(DepartmentName, Minority.x, White.y) %>%
  gather("Searched", "Percent", 2:3)  


stops_search2 <- single_all %>%
  select(DepartmentName, Minority.x, White.y) %>%
  gather("Searched", "Percent", 2:3)  


stops_search2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_search2$DepartmentName)

stops_search3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, Minority.x, White.y) %>%
  gather("Searched", "Percent", 2:3)  

names(stops_search3)[names(stops_search3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'

stops_search <- rbind(stops_search1, stops_search2)
stops_search <- rbind(stops_search, stops_search3)


stops_search$Searched <- gsub(".x", "", stops_search$Searched)
stops_search$Searched <- gsub(".y", "", stops_search$Searched)

ggplot(stops_search, aes(Searched, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Who was searched") 

```



---

## Residential stops

For officer `r single_officer$ReportingOfficerIdentificationID[1]`

* **`r single_officer$resident_p[1]`** percent were residents
* **`r single_officer$not.resident_p[1]`** were non-residents


In `r single_dept$DepartmentName[1]`

* **`r single_dept$resident_p[1]`** percent were residents
* **`r single_dept$not.resident_p[1]`** were non-residents

Statewide

* **`r single_all$resident_p[1]`** percent were men
* **`r single_all$not.resident_p[1]`** were women

---

## Black drivers: Of those stopped, how many were residents or non-residents?



```{r section7, fig.width=8, fig.height=4, echo=FALSE}

stops_res_min1 <- single_dept %>%
  select(DepartmentName, not.resident_b_p, resident_b_p) %>%
  gather("Status", "Percent", 2:3)  


stops_res_min2 <- single_all %>%
  select(DepartmentName, not.resident_b_p, resident_b_p) %>%
  gather("Status", "Percent", 2:3)  


stops_res_min2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_res_min2$DepartmentName)

stops_res_min3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, not.resident_b_p, resident_b_p) %>%
  gather("Status", "Percent", 2:3)  


names(stops_res_min3)[names(stops_res_min3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'

stops_res_min <- rbind(stops_res_min1, stops_res_min2)
stops_res_min <- rbind(stops_res_min, stops_res_min3)

stops_res_min$Status <- gsub("_p", "", stops_res_min$Status)

ggplot(stops_res_min, aes(Status, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Black residents pulled over") 

```


---

## Hispanic drivers: Of those stopped, how many were residents or non-residents?

```{r section8, fig.width=8, fig.height=4, echo=FALSE}

stops_res_min1 <- single_dept %>%
  select(DepartmentName, not.resident_h_p, resident_h_p) %>%
  gather("Status", "Percent", 2:3)  


stops_res_min2 <- single_all %>%
  select(DepartmentName, not.resident_h_p, resident_h_p) %>%
  gather("Status", "Percent", 2:3)  


stops_res_min2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_res_min2$DepartmentName)


stops_res_min3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, not.resident_h_p, resident_h_p) %>%
  gather("Status", "Percent", 2:3)  

names(stops_res_min3)[names(stops_res_min3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'

stops_res_min <- rbind(stops_res_min1, stops_res_min2)
stops_res_min <- rbind(stops_res_min, stops_res_min3)

stops_res_min$Status <- gsub("_p", "", stops_res_min$Status)

ggplot(stops_res_min, aes(Status, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Hispanic residents pulled over") 

```



---

## Non-white drivers: Of those stopped, how many were residents or non-residents?

```{r section9, fig.width=8, fig.height=4, echo=FALSE}

stops_res_min1 <- single_dept %>%
  select(DepartmentName, not.resident_m_p, resident_m_p) %>%
  gather("Status", "Percent", 2:3)  


stops_res_min2 <- single_all %>%
  select(DepartmentName, not.resident_m_p, resident_m_p) %>%
  gather("Status", "Percent", 2:3)  


stops_res_min2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_res_min2$DepartmentName)

stops_res_min3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, not.resident_m_p, resident_m_p) %>%
  gather("Status", "Percent", 2:3)  

names(stops_res_min3)[names(stops_res_min3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'


stops_res_min <- rbind(stops_res_min1, stops_res_min2)
stops_res_min <- rbind(stops_res_min, stops_res_min3)

stops_res_min$Status <- gsub("_p", "", stops_res_min$Status)

ggplot(stops_res_min, aes(Status, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Minority residents pulled over") 

```


---

## Type of stops

```{r section10, fig.width=8, fig.height=4, echo=FALSE}

stops_ser_min1 <- single_dept %>%
  select(DepartmentName, general.enforcement_p, blind.enforcement_p, spot.check_p) %>%
  gather("Stops", "Percent", 2:4)  


stops_ser_min2 <- single_all %>%
  select(DepartmentName, general.enforcement_p, blind.enforcement_p, spot.check_p) %>%
  gather("Stops", "Percent", 2:4)  


stops_ser_min2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_ser_min2$DepartmentName)


stops_ser_min3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, general.enforcement_p, blind.enforcement_p, spot.check_p) %>%
  gather("Stops", "Percent", 2:4)  

names(stops_ser_min3)[names(stops_ser_min3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'

stops_ser_min <- rbind(stops_ser_min1, stops_ser_min2)
stops_ser_min <- rbind(stops_ser_min, stops_ser_min3)

stops_ser_min$Stops <- gsub("_p", "", stops_ser_min$Stops)

ggplot(stops_ser_min, aes(Stops, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Percent of type of stops out of all stops") 


```



## Spot checks

```{r section11, fig.width=8, fig.height=4, echo=FALSE}

stops_ser_min1 <- single_dept %>%
  select(DepartmentName, Minority.y, White.x.1) %>%
  gather("Stops", "Percent", 2:3)  


stops_ser_min2 <- single_all %>%
  select(DepartmentName, Minority.y, White.x.1) %>%
  gather("Stops", "Percent", 2:3)  



stops_ser_min2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_ser_min2$DepartmentName)

stops_ser_min3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, Minority.y, White.x.1) %>%
  gather("Stops", "Percent", 2:3)  

names(stops_ser_min3)[names(stops_ser_min3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'

stops_ser_min <- rbind(stops_ser_min1, stops_ser_min2)
stops_ser_min <- rbind(stops_ser_min, stops_ser_min3)


stops_ser_min$Stops <- gsub("_p", "", stops_ser_min$Stops)

ggplot(stops_ser_min, aes(Stops, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Percent of spot checks for minorities out of total stops") 

```


## Blind enforcement

```{r section12, fig.width=8, fig.height=4, echo=FALSE}

stops_ser_min1 <- single_dept %>%
  select(DepartmentName, Minority, White.y.1) %>%
  gather("Stops", "Percent", 2:3)  


stops_ser_min2 <- single_all %>%
  select(DepartmentName, Minority, White.y.1) %>%
  gather("Stops", "Percent", 2:3)  



stops_ser_min2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_ser_min2$DepartmentName)

stops_ser_min3 <- single_officer %>%
  select(ReportingOfficerIdentificationID, Minority, White.y.1) %>%
  gather("Stops", "Percent", 2:3)  

names(stops_ser_min3)[names(stops_ser_min3) == 'ReportingOfficerIdentificationID'] <- 'DepartmentName'

stops_ser_min <- rbind(stops_ser_min1, stops_ser_min2)
stops_ser_min <- rbind(stops_ser_min, stops_ser_min3)

stops_ser_min$Stops <- gsub("_p", "", stops_ser_min$Stops)

ggplot(stops_ser_min, aes(Stops, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), vjust = -0.5, size=3) + labs(title="Percent of blind enforcement for minorities out of total stops") 

```

## About `r single_dept$DepartmentName[1]`

* Number of employees in department: **`r single_dept$total_leo[1]`** (`r single_dept$leo_per_capita[1]` per 10,000 residents)
* Number of officers: **`r single_dept$total_officers[1]`** (`r single_dept$officers_per_capita[1]` per 10,000 residents)
* Number of civilian workers: **`r single_dept$total_civilians[1]`** (`r single_dept$civ_per_capita[1]` per 10,000 residents)



```{r section13, fig.width=8, fig.height=8, echo=FALSE}

stops_reason1 <- single_dept %>%
  select(DepartmentName, violent_crime_pc, murder_manslaughter_pc, rape_pc, robbery_pc, aggravated_robbery_pc, property_crime_pc, burglary_pc, larceny_theft_pc, motor_vehicle_theft_pc, arson_pc) %>%
  gather("Crime", "Percent", 2:11)  

stops_reason2 <- single_all %>%  
  select(DepartmentName, violent_crime_pc, murder_manslaughter_pc, rape_pc, robbery_pc, aggravated_robbery_pc, property_crime_pc, burglary_pc, larceny_theft_pc, motor_vehicle_theft_pc, arson_pc) %>%
  gather("Crime", "Percent", 2:11)  

stops_reason2$DepartmentName <- gsub("Connecticut average", "Connecticut", stops_reason2$DepartmentName)

stops_reason <- rbind(stops_reason1, stops_reason2)
stops_reason$Crime <- gsub("_pc", "", stops_reason$Crime)

ggplot(stops_reason, aes(Crime, Percent, fill = DepartmentName, label = Percent)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(position = position_dodge(0.9), hjust = -0.2, size=3) + labs(title="Crime per 10,000 residents") + coord_flip()
```

