---
title: "Counties housing and metro rent analysis"
author: "Andrew Ba Tran"
date: "May 31, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is the methodology used behind the story: [Only California and Hawaii beat some metros in CT for price of a two-bedroom](http://trendct.org/2016/06/01/comparing-the-cost-of-renting-a-two-bedroom/).

Visit the [repo](https://github.com/trendct/data/tree/master/2016/05/housing-income-oor) for the data used in this analysis or visit the National Low Income Housing Coalition's [Out of Reach 2016](http://nlihc.org/oor) page for even more thorough data sets.

A map used in this analysis is based on code from [rud.is](https://rud.is/b/2016/03/29/easier-composite-u-s-choropleths-with-albersusa/).

What this analysis does:

* Download and bring in 50 spreadsheets for each state from NLIHC.
* Create dataframes specific to what we're analyzing
* Generate map visualizations of the data
* Create a chart illustrating the gap between wage data
* Create a datatable of all metro areas and cost of two-bedroom dwelling and necessary income


```{r libraries, warning=FALSE, message=FALSE}
library(RCurl)
library(readxl)
library(dplyr)
library(stringr)
library(tidyr)
library(rgeos)
library(maptools)
library(ggplot2)   # devtools::install_github("hadley/ggplot2") only if you want subtitles/captions
library(ggalt)
library(ggthemes)
library(albersusa) # devtools::install_github("hrbrmstr/albersusa")
library(viridis)
library(scales)
library(DT)
```

```{r bringing_in_data, warning=FALSE, message=FALSE, include=FALSE}


for (i in 1:length(state.abb)) {
  file_name <- paste0("2016-OOR-", state.abb[i], ".xls")
  file_url <- paste0("http://nlihc.org/sites/default/files/oor/files/reports/state/", file_name)
  
  download.file(file_url, paste0("data/", file_name), mode="wb")
  state <- read_excel(paste0("data/", file_name), sheet=2)

  
  state_only <- subset(state, TYPE=="METRO")
  counties <- subset(state, TYPE=="COUNTY")
  
  if (i==1) {
    states_all <- state_only
    counties_all <- counties
    
  } else {
    states_all <- rbind(states_all, state_only)
    counties_all <- rbind(counties_all, counties)
    
  }
}
```

```{r restructring_data, warning=FALSE, message=FALSE}
colnames(states_all) <- make.names(colnames(states_all))
colnames(counties_all) <- make.names(colnames(counties_all))

states_all <- states_all[c("STNAME", "COUNTY.METRO", "Two.bedroom.FMR", "Income.needed.to.afford.2.bdrm.FMR")]
counties_all <- counties_all[c("STNAME", "COUNTY.METRO", "Two.bedroom.FMR", "Income.needed.to.afford.2.bdrm.FMR")]

colnames(states_all) <- c("State", "metro", "two.bed.cost", "income.two.bed")
colnames(counties_all) <- c("State", "county", "two.bed.cost", "income.two.bed")
```

```{r preparing_data_for_mapping, warning=FALSE, message=FALSE}

counties_all$county_name <- paste(counties_all$county, counties_all$State)

cmap <- fortify(counties_composite(), region="fips")

cmap$state_id <- substr(cmap$id, 0,2)
substr("abcdef", 2, 4)

cmap2 <-fortify(counties_composite(), region="state") 

county_fix <- read.csv("data/county_fips.csv", stringsAsFactors=FALSE)
county_fix$X <- NULL

county_fix$fips <- as.character(county_fix$fips)

county_fix$fips <- ifelse(nchar(county_fix$fips)==4, paste0("0", county_fix$fips), county_fix$fips)
county_fix$polyname <- gsub("\\.", "", county_fix$polyname)


counties_all$fips_fix <- gsub(" County", "", counties_all$county)
counties_all$fips_fix <- paste0(counties_all$State, ",", counties_all$fips_fix)
counties_all$fips_fix <- str_to_lower(counties_all$fips_fix)
counties_all$fips_fix <- gsub("\\.", "", counties_all$fips_fix)
counties_all$fips_fix <- gsub("\\*", "", counties_all$fips_fix)
counties_all$fips_fix <- gsub("\\†", "", counties_all$fips_fix)
counties_all$fips_fix <- gsub(" parish", "", counties_all$fips_fix)
counties_all$fips_fix <- gsub(" city", "", counties_all$fips_fix)
counties_all$fips_fix <- gsub("dekalb", "de kalb", counties_all$fips_fix)
counties_all$fips_fix <- gsub("\\'", "", counties_all$fips_fix)
counties_all$fips_fix <- gsub("desoto", "de soto", counties_all$fips_fix)
counties_all$fips_fix <- gsub("dupage", "du page", counties_all$fips_fix)
counties_all$fips_fix <- gsub("laporte", "la porte", counties_all$fips_fix)
counties_all$fips_fix <- gsub("louisiana,st martin", "louisiana,st martin:north", counties_all$fips_fix)
counties_all$fips_fix <- gsub("washington,pierce", "washington,pierce:main", counties_all$fips_fix)
counties_all$fips_fix <- gsub("washington,san juan", "washington,san juan:lopez island", counties_all$fips_fix)
counties_all$fips_fix <- gsub("galveston", "galveston:main", counties_all$fips_fix)
counties_all$fips_fix <- gsub("dewitt", "de witt", counties_all$fips_fix)
counties_all$fips_fix <- gsub("lamoure", "la moure", counties_all$fips_fix)
counties_all$fips_fix <- gsub("currituck", "currituck:knotts", counties_all$fips_fix)
counties_all$fips_fix <- gsub("nevada,carson", "nevada,carson city", counties_all$fips_fix)
counties_all$fips_fix <- gsub("okaloosa", "okaloosa:main", counties_all$fips_fix)


counties_all$fips_fix <- str_trim(counties_all$fips_fix)

colnames(county_fix) <- c("fips", "fips_fix")

hiak <- data.frame(fips=c("02013", "02016", "02020", "02050", "02060", "02068", "02070", "02090", "02100", "02110", "02122", "02130", "02150", "02164", "02170", "02180", "02185", "02188", "02201", "02220", "15001", "15003", "15005", "15007", "15009", "02232", "02240", "02261", "02270", "02280", "02282", "02290", "02105", "02280", "02280"),
                   fips_fix=c("alaska,aleutians east borough", "alaska,aleutians west census area", "alaska,anchorage municipality", "alaska,bethel census area", "alaska,bristol bay borough", "alaska,denali borough", "alaska,dillingham census area", "alaska,fairbanks north star borough", "alaska,haines borough", "alaska,juneau and borough", "alaska,kenai peninsula borough", "alaska,ketchikan gateway borough", "alaska,kodiak island borough", "alaska,lake and peninsula borough", "alaska,matanuska-susitna borough", "alaska,nome census area", "alaska,north slope borough", "alaska,northwest arctic borough", "alaska,prince of wales-hyder census area", "alaska,sitka and borough", "hawaii,hawaii", "hawaii,honolulu", "hawaii,kalawao", "hawaii,kauai", "hawaii,maui", "alaska,skagway municipality", "alaska,southeast fairbanks census area", "alaska,valdez-cordova census area", "alaska,wade hampton census area", "alaska,wrangell-petersburg census area", "alaska,yakutat and borough", "alaska,yukon-koyukuk census area", "alaska,hoonah-angoon census area", "alaska,wrangell and borough", "alaska,petersburg census area"))

county_fix <- rbind(county_fix, hiak)

counties_all <- left_join(counties_all, county_fix)

```

# Map

```{r map1, fig.width=10, fig.height=7}

gg <- ggplot()
gg <- gg + geom_map(data=cmap, map=cmap,
                    aes(x=long, y=lat, map_id=id),
                    color="#2b2b2b", size=0.05, fill=NA)
gg <- gg + geom_map(data=counties_all, map=cmap,
                    aes(fill=two.bed.cost, map_id=fips),
                    color="#2b2b2b", size=0.05)
gg <- gg + scale_fill_gradient(name="Cost",labels=dollar_format(), low = "#ce1256", high = "#4eb3d3")
gg <- gg + coord_proj(us_laea_proj)
gg <- gg + labs(title="Average monthly cost of a two-bedroom dwelling",
                subtitle="Data from some counties, mostly in New England, were not available",
                caption="SOURCE: National Low Income Housing Coalition\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_map(base_family="Arial Narrow")
gg <- gg + theme(legend.position=c(0.8, 0.25))
gg <- gg + theme(plot.title=element_text(face="bold", size=14, margin=margin(b=6)))
gg <- gg + theme(plot.subtitle=element_text(size=10, margin=margin(b=-14)))
gg
```

# Connecticut only

```{r CT_only, warning=FALSE, message=FALSE, fig.height=6, fig.width=9}

ct_only <- subset(states_all, State=="Connecticut")

ct_only$State <- NULL
ct_only$income.two.bed <- NULL
ct_only$metro <- gsub(" HMFA", "", ct_only$metro)
ct_only$metro <- gsub("\\*", "", ct_only$metro)
ct_only$metro <- str_trim(ct_only$metro)

g <- ggplot(ct_only, aes(metro, two.bed.cost))
g <- g + geom_bar(stat="identity") 
g <- g + coord_flip() 
g <- g + labs(title="Monthly average rent for CT 2-bedroom")
g <- g + theme_minimal()

g
```

```{r datatable, fig.height=7, fig.width=10}
states_all$two.bed.cost <- round(states_all$two.bed.cost, 2)
states_all$metro <- gsub(" HMFA", "", states_all$metro)
states_all$metro <- gsub(" MSA", "", states_all$metro)
states_all$metro <- gsub("\\*", "", states_all$metro)

states_all$rank <- rank(-states_all$two.bed.cost, ties.method= "max")
datatable(states_all)
```